---
title: "Gian Mella Check Up Memo"
subtitle: "Final Project (302)" 
author: "Gian Mella"
date: November 14, 2022
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
    fig-dpi: 300
execute:
  warning: false
from: markdown+emoji 
editor_options: 
  chunk_output_type: console
---

## Data Source

My primary data source is from Stathead: [https://stathead.com/tiny/hXqZ6](https://stathead.com/tiny/hXqZ6).

This dataset lists every NBA player who played over 10 minutes per game in last year's playoffs (2021-22 NBA season), sorted by descending win shares. This table shows a wide variety of data for each player, such as per game stats, shooting percentages, and some advanced statistics as well. There are 152 players in the dataset.

## Demonstration of Progress

```{r}

# loading packages

library(tidyverse)
library(janitor)
library(skimr)

bball_data <- read_csv("data/unprocessed/Bball_data.txt") %>%
  clean_names()

skim_without_charts(bball_data)
```

I was able to successfully load in the my dataset and perfomed a quick skim of the data. To my relief, missingness was very minimal and easy to explain. But I knew I had to fix a few things with my data before I could proceed into making visualizations. 

First, I got rid of of columns in the dataset that I felt were unnecessary, such as the age of the player, what season it is (since it's the same for everyone), and some duplicate columns for minutes played and win shares. 

```{r}
bball_data <-
  bball_data %>%
  select(-c(rk, ws_10, season, mp_16, x9999, age))
```


I also needed to change a lot of my column names. There are many columns that wouldn't make sense when the average person is looking at my data, such as variables "g", "x3pa", or "x2p_percent". I renamed those columns along with some others for increased clarity.

```{r}
bball_data <- rename(bball_data, `win_shares` = `ws_3`, `mins` = `mp_4`, `games_played` = `g`, `games_started` = `gs`, `2p` = `x2p`, `2pa` = `x2pa`, `3p` = `x3p`, `3pa` = `x3pa`, `fouls` = `pf`, `2p_percent` = `x2p_percent`, `3p_percent` = `x3p_percent`, `efg_percent` = `e_fg_percent`, `position` = `pos`)
```

In addition, for all the shooting percentage variables, I noticed that they were in terms of decimals and not shooting percentages. For example a 51.3% mark from 3pt range is listed as ".513". I'd rather have it as a percentage, so I converted all of those shooting decimals to percentages by multiplying by 100. 

```{r}
bball_data1 <- bball_data %>%
  mutate(
  `fg_percent` = `fg_percent` * 100,
  `2p_percent` = `2p_percent` * 100,
  `3p_percent` = `3p_percent` * 100,
  `ft_percent` = `ft_percent` * 100,
  `ts_percent` = `ts_percent` * 100,
  `efg_percent` = `efg_percent` * 100
  )
```

Lastly, I'm going create a little visualization that gives me a glimpse into Jimmy Butler's overall playoff performance last year versus everyone else. The following chart shows Butler's points per game versus the league average.

```{r}
(jimmy_comparison <- bball_data1 %>%
  mutate(player = ifelse(player == "Jimmy Butler", "Jimmy Butler", "All Other Players")) %>%
  group_by(player) %>%
  summarize(`Points Per Game (2021-22 Playoffs)` = mean(`pts`))
)

jimmy_comparison %>%
  ggplot(aes(player, `Points Per Game (2021-22 Playoffs)`)) +
  geom_col(fill = "light blue")
```

## Assessment of Progress

I am pretty satisfied with the progress I have made so far. Loading the data was pretty simple, and I didn't run into any major problems  with my data that stalled my work. Every fix that I wanted to make with my dataset I was able to do. In the time that I've spent exploring and working with the data, I've already developed some ideas for visualizations I can possibly make. I think I am on track with this project and actually hope to start working on it earlier and potentially submit it by the early deadline.
